\section{Integral Domains, Fields of Fractions, Prime Ideals}

Notice that in the ring \(\ZZ / 6\), we have \(2 \cdot 3 = 0\) but \(2 \neq 0\) and \(3 \neq 0\).

\begin{definition}[Integral Domain, Zero Divisor]
    \label{def:integral-domain-zero-divisor}%
    A non-zero ring \(R\) is an \emph{integral domain}, if for all \(a, b \in R\), we have
    \[
        a b = 0 \implies a = 0 \quad\text{or}\quad b = 0.
    \]

    A \emph{zero divisor} if a non-zero \(a\), such that there exists \(b \neq 0\) if \(ab = 0\).
\end{definition}

For example, in \(\ZZ / 6\), we say both \(2\) and \(3\) are zero divisors.

\begin{examples}[Examples of Integral Domains]
    \label{ex:integral-domain}%
    \begin{itemize}
        \item Any field is an integral domain. Indeed, if \(a b = 0\) in a field, say \(b \neq 0\), then \(b\inverse\) exists, and
    \[
        a = a b b\inverse = a \cdot 0 = 0
    \]
    so \(a = 0\).
        \item Similarly, any subring of a field is also an integral domain. This is really how we should think of integral domains.
        \item The rings \(\ZZ\), \(\ZZ\brac{i}\), \(\QQ\), \(\CC\), \(\RR\brac{x}\) \(\ZZ\brac{X}\) are integral domains.
        \item The rings \(\ZZ / 6\), \(\ZZ / \para{pq}\), \(\ZZ / 2 \times \ZZ / 2\), etc.
        \item In particular, direct products cannot be integral domains, since \(\para{0, 1}\) and \(\para{1, 0}\) product to \(0\).
    \end{itemize}
\end{examples}

\begin{lemma}[Finite Integral Domains are Finite Fields]
    \label{lem:finite-integral-domain-finate-field}%
    Let \(R\) be a finite integral domain. Then \(R\) is a field.
\end{lemma}

An example of this is \(\ZZ / p\) for a prime \(p\).

\begin{proof}
    Let \(a \in R\) be non-zero, we want to prove it is a unit. Consider the multiplication function
    \[
        \function{\mu_a}{R}{R}{r}{ar.}
    \]

    This is a group homomorphism for the underlying group structure. Since \(R\) is an integral domain, by definition, we have
    \[
        \ker \para{\mu_a} = \brce{0}.
    \]

    So \(\mu_a\) is injective. Recall that \(R\) is finite, which means \(\mu_a\) is bijective.

    In particular, \(1 \in \img \para{\mu_a}\), so there is some \(b \in R\) with \(b \neq 0\), such that
    \[
        a \cdot b = 1
    \]
    so \(a\) is indeed a unit, finishing our proof.
\end{proof}

\begin{definition}[Field of Fractions]
    \label{def:field-of-fractions}%
    Let \(R\) be an integral domain. A \emph{field of fractions} of \(R\) is a field \(F\), such that
    \begin{enumerate}
        \item \(R \subring F\) is a subring, meaning there is an injective ring homomorphism from \(R\) to \(F\); and
        \item any \(x \in F\) can be written as \(a b \inverse\), for \(a, b \in R\), and \(b\inverse\) as the multiplicative inverse of \(b\) in \(F\).
    \end{enumerate}
\end{definition}

The first requirement is `sticking \(R\) into some field \(F\)', while the second requirement is `\(F\) is only contained of fractions in \(R\)'.

\begin{example}[Field of Fractions]
    \label{ex:field-of-fractions}
    The field \(\QQ\) is a field of fractions of \(\ZZ\).
\end{example}

\begin{theorem}[Existence of Field of Fractions]
    \label{thm:existence-field-of-fractions}%
    Every integral domain \(R\) has a field of fractions.
\end{theorem}

Note \autoref{thm:existence-field-of-fractions} tells us that all integral rings are effectively subrings. The proof is inspired by the construction of \(\QQ\) by \(\ZZ\).

\begin{proof}
    Define a set
    \[
        S = \set{\para{a, b} \in R \times R}{b \neq 0}.
    \]

    We define a relation \(\sim\) over \(S\), being
    \[
        \para{a, b} \sim \para{c, d} \iff ad = bc,
    \]
    where here the multiplication is in \(R\).

    We want to show \(\sim\) is an equivalence relation. \emph{Symmetry} and \emph{reflective} are obvious. We conduct a non-trivial check on the \emph{transitivity}. Suppose that
    \[
        \para{a, b} \sim \para{c, d} \quad\text{and}\quad \para{c, d} \sim \para{e, f}.
    \]

    We have \(ad = bc\) and \(cf = de\). Multiplying the first equality by \(f\), and the second by \(b\), we have
    \[
        adf = bcf \quad\text{and}\quad bcf = bde,
    \]
    and hence \(adf = bde\). So we may rearrange and combine terms to get
    \[
        d \para{af - be} = 0.
    \]

    We have \(d \neq 0\), by our definition of \(S\). Using that \(R\) is an integral domain, we have \(af - be = 0\), so indeed \(af = be\), and we have \(\para{a, b} \sim \para{e, f}\).

    Now, we take the equivalence classes
    \[
        F = S / \sim,
    \]
    and we denote \(\frac{a}{b} = \brac{\para{a, b}}\) for the conjugacy class. We make \(F\) a ring by the following operations.

    \[
        \frac{a}{b} + \frac{c}{d} = \frac{ad + bc}{bd} \quad\text{and}\quad \frac{a}{b} \cdot \frac{c}{d} = \frac{ac}{bd}.
    \]

    Elementary checks show that these operations are well-defined and \(F\) is a ring.

    To see \(F\) is a field, note that \(\frac{a}{b} \neq 0_F\), then \(\frac{a}{b} \neq \frac{0}{1}\), so \(a \cdot 1 \neq b \cdot 0\), so \(a \neq 0\). Then \(\frac{b}{a} \in F\) is an inverse of \(\frac{a}{b}\).

    Finally, we define a homomorphism
    \[
        \function{\phi}{R}{F}{r}{\frac{r}{1}.}
    \]

    Clearly, \(\phi\) is a ring homomorphism, and observe that
    \[
        \ker\para{\phi} = \set{r \in R}{\frac{r}{1} = 0} = \brce{0}.
    \]

    By \autoref{thm:first-isomorphism-theorem-ring} (First Isomorphism Theorem for Rings), we have \(\img \para{\phi} \isomorphic R\). Since every \(\frac{a}{b} \in R\) can be written as
    \[
        \frac{a}{b} = \frac{a}{1} \cdot \frac{1}{b} = \para{\frac{a}{1}} \cdot \para{\frac{b}{1}}\inverse,
    \]
    so we have completed the proof.
\end{proof}